---
title: "Bumble bee nesting & ovary development"
author: "ES"
date: "7/23/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# knitr::opts_chunk$set(fig.width=3.5, fig.height=4) 
    # use only to create vertical vos 2-panel plots

# knitr::opts_chunk$set(fig.width=3, fig.height=4) 
    # use only to create vertical ALL 2-panel plots

```

# load packages
```{r}
library(tidyverse)
library(ggpubr)   # ggdensity
library(car)      # qqp
library(lme4)     # glmer
library(sjPlot)   # tab_model
library(effects)  # plot(allEffects)
library(multcomp) # posthoc tukey
library(MuMIn)    # model.sel
```

# load data
```{r}
fq <- read.csv("./data/raw_data.csv", header = TRUE, 
               na.strings = c("na", "NA" , "", " "))

# remove stationary queens (outside of 3 focal behavioral categories)
fq <- fq[fq$stationary == 0,]
fq
```

# format fieldqueen data
fq has one line per queen (used in max oocylte length & resorption analyses)
```{r}
# BEHAVIOR
# add column for behavior to put all behaviors into a single column
fq$behav <- NA
# pollen collecting / nest searching /  nectaring
for (i in 1:nrow(fq)) {
  if (fq$pollen[i] == 1) {
    fq$behav[i] <- "pollen"
  }
  if (fq$ns[i] == 1) {
    fq$behav[i] <- "nestsearch"
  }
  if (fq$nectaring[i] == 1) {
    fq$behav[i] <- "nectaring"
  }
}

# code columns as correct type
fq$pollen <- as.character(fq$pollen)
fq$ns <- as.character(fq$ns)
fq$nectaring <- as.character(fq$nectaring)
fq$mites <- as.character(fq$mites)
fq$sphaer <- as.character(fq$sphaer)
fq$otherMacroPara <- as.character(fq$otherMacroPara)
fq$nosema <- as.character(fq$nosema)
fq$apocys <- as.character(fq$apocys)
fq$otherMicroPara <- as.character(fq$otherMicroPara)

## BODY SIZE
# average wing cells together for proxy for body size
fq$bodySize <- rowMeans(fq[, c("wing1", "wing2")], na.rm = TRUE)
  
## DATE
# format collection date as date
fq$date <- as.character(fq$date)
fq$date <- as.Date(fq$date, "%Y%m%d")
fq

## SUMMARIZE OVARY DATA
# convert oocyte length measurements to mm 
# (CA bees: 5 eyepiece units = 4.1mm // 1 ocular micrometer unit = 0.82mm/unit)
# (ME bees: 10 eyepiece units = 4mm // 1 ocular micrometer = 0.4mm)
for (i in 1:nrow(fq)) {
  if (startsWith(fq$beeID[i], "CA")) {
    fq$o1l[i] <- fq$o1l[i] * 0.82
    fq$o2l[i] <- fq$o2l[i] * 0.82
    fq$o3l[i] <- fq$o3l[i] * 0.82
    fq$o4l[i] <- fq$o4l[i] * 0.82
    fq$o5l[i] <- fq$o5l[i] * 0.82
    fq$o6l[i] <- fq$o6l[i] * 0.82
    fq$o7l[i] <- fq$o7l[i] * 0.82
    fq$o8l[i] <- fq$o8l[i] * 0.82
  }
  if (startsWith(fq$beeID[i], "ME")) {
    fq$o1l[i] <- fq$o1l[i] * 0.4
    fq$o2l[i] <- fq$o2l[i] * 0.4
    fq$o3l[i] <- fq$o3l[i] * 0.4
    fq$o4l[i] <- fq$o4l[i] * 0.4
    fq$o5l[i] <- fq$o5l[i] * 0.4
    fq$o6l[i] <- fq$o6l[i] * 0.4
    fq$o7l[i] <- fq$o7l[i] * 0.4
    fq$o8l[i] <- fq$o8l[i] * 0.4
  }
}

# count number of resorbed oocytes per queen & number of na oocytes per queen
fq$oocNres <- NA
fq$oocNNA <- NA

for (i in 1:nrow(fq)) {
  nR <- 0
  nNA <- 0
  if (is.na(fq$o1s[i])) {
    nNA <- nNA + 1
  } else if (fq$o1s[i] == "R") {
    nR <- nR + 1
  }
  if (is.na(fq$o2s[i])) {
    nNA <- nNA + 1
  } else if (fq$o2s[i] == "R") {
    nR <- nR + 1
  }
  if (is.na(fq$o3s[i])) {
    nNA <- nNA + 1
  } else if (fq$o3s[i] == "R") {
    nR <- nR + 1
  }
  if (is.na(fq$o4s[i])) {
    nNA <- nNA + 1
  } else if (fq$o4s[i] == "R") {
    nR <- nR + 1
  }
  if (is.na(fq$o5s[i])) {
    nNA <- nNA + 1
  } else if (fq$o5s[i] == "R") {
    nR <- nR + 1
  }
  if (is.na(fq$o6s[i])) {
    nNA <- nNA + 1
  } else if (fq$o6s[i] == "R") {
    nR <- nR + 1
  }
  if (is.na(fq$o7s[i])) {
    nNA <- nNA + 1
  } else if (fq$o7s[i] == "R") {
    nR <- nR + 1
  }
  if (is.na(fq$o8s[i])) {
    nNA <- nNA + 1
  } else if (fq$o8s[i] == "R") {
    nR <- nR + 1
  }
  fq$oocNres[i] <- nR
  fq$oocNNA[i] <- nNA
}

# calculate the proportion of oocytes resorbed per queen
fq$oocPropR <- fq$oocNres / (8 - fq$oocNNA)
```

# format fieldovary data 
fo has one line per ovary per queen (used in oocyte length analyses)
```{r}
## MAKE SEPARATE DF OF ALL OOCYTE LENGTHS (including resorbed oocytes)
# gather oocyte lengths for graphing together
lendf <- fq %>% gather(o1l, o2l, o3l, o4l, o5l, o6l, o7l, o8l,
                         key = "oocyte",
                         value = "oocLen")

## MAKE SEPARATE DF OF OOCYTE LENGTH FOR UNRESORBED OOCYTES ONLY
# remove resorbed oocytes
lenUnRdf <- lendf
lenUnRdf$oocLen[lenUnRdf$oocyte == "o1l" & lenUnRdf$o1s == "R"] <- NA
lenUnRdf$oocLen[lenUnRdf$oocyte == "o2l" & lenUnRdf$o2s == "R"] <- NA
lenUnRdf$oocLen[lenUnRdf$oocyte == "o3l" & lenUnRdf$o3s == "R"] <- NA
lenUnRdf$oocLen[lenUnRdf$oocyte == "o4l" & lenUnRdf$o4s == "R"] <- NA
lenUnRdf$oocLen[lenUnRdf$oocyte == "o5l" & lenUnRdf$o5s == "R"] <- NA
lenUnRdf$oocLen[lenUnRdf$oocyte == "o6l" & lenUnRdf$o6s == "R"] <- NA
lenUnRdf$oocLen[lenUnRdf$oocyte == "o7l" & lenUnRdf$o7s == "R"] <- NA
lenUnRdf$oocLen[lenUnRdf$oocyte == "o8l" & lenUnRdf$o8s == "R"] <- NA
#rename the column to differentiate from all oocyte lengths
names(lenUnRdf)[names(lenUnRdf) == "oocLen"] <- "oocLen_unR"
lenUnRdf


## MAKE SEPARATE OOCYTE STAGE DF
# gather oocyte stages
stagedf <- fq %>% gather(o1s, o2s, o3s, o4s, o5s, o6s, o7s, o8s,
                         key = "oocyte",
                         value = "stage")
# mark resorbed oocytes in own column
stagedf$res <- 0
# if stage is R, make res 1
stagedf$res[stagedf$stage == "R"] <- 1
# if stage is NA, make res NA
stagedf$res[is.na(stagedf$stage)] <- NA
# make resorbed oocytes NA stage 
stagedf$stage[stagedf$stage == "R"] <- 4
stagedf$stage <- as.numeric(stagedf$stage)
stagedf

# MERGE stagedf, lenUnRdf, & lendf TO MAKE ONE DF w/ 1 ROW PER OVARY PER QUEEN
fo <- cbind(lendf, lenUnRdf$oocLen_unR, stagedf$stage, stagedf$res)
# rename added columns
names(fo)[names(fo) == "lenUnRdf$oocLen_unR"] <- "oocLen_unR"
names(fo)[names(fo) == "stagedf$stage"] <- "stage"
names(fo)[names(fo) == "stagedf$res"] <- "res"
fo

## GET SUMMARY STATS & ADD THEM BACK INTO fq
# pivot wider fo for summary stats
sumdf <- pivot_wider(fo, names_from = oocyte, 
                     values_from = c(oocLen, oocLen_unR, stage, res, oocPropR),
                     id_col = c(beeID)) 
# calculate max unresorbed oocyte length for each queen
maxUnR_Lendf <- sumdf[, c("oocLen_unR_o1l", "oocLen_unR_o2l", 
                          "oocLen_unR_o3l", "oocLen_unR_o4l", 
                          "oocLen_unR_o5l", "oocLen_unR_o6l", 
                          "oocLen_unR_o7l", "oocLen_unR_o8l")]
max <- apply(maxUnR_Lendf, 1, max, na.rm = TRUE) 
max[is.infinite(max)] <- NA
fq$oocMaxUnR_Len <- max
# calculate min unresorbed oocyte length for each queen
maxUnR_Lendf <- sumdf[, c("oocLen_unR_o1l", "oocLen_unR_o2l", 
                          "oocLen_unR_o3l", "oocLen_unR_o4l", 
                          "oocLen_unR_o5l", "oocLen_unR_o6l", 
                          "oocLen_unR_o7l", "oocLen_unR_o8l")]
min <- apply(maxUnR_Lendf, 1, min, na.rm = TRUE) 
min[is.infinite(min)] <- NA
fq$oocMinUnR_Len <- min
# calculate mean unresorbed oocyte length for each queen
fq$oocMeanUnR_Len <- rowMeans(sumdf[, c("oocLen_unR_o1l", "oocLen_unR_o2l", 
                                        "oocLen_unR_o3l", "oocLen_unR_o4l", 
                                        "oocLen_unR_o5l", "oocLen_unR_o6l", 
                                        "oocLen_unR_o7l", "oocLen_unR_o8l")], 
                              na.rm = TRUE)
fq
```

# summarize
```{r}
# number of bees in each behavior by species

# vos
print(paste(nrow(fq[fq$species == "vosnesenskii" & fq$behav == "nestsearch", ]),
            "vosnesenskii nest searching"))
print(paste(nrow(fq[fq$species == "vosnesenskii" & fq$behav == "pollen", ]), 
            "vosnesenskii pollen collecting"))
print(paste(nrow(fq[fq$species == "vosnesenskii" & fq$behav == "nectaring", ]), 
            "vosnesenskii nectaring"))

# substantial mites
paste(nrow(fq[fq$mites == "1" & fq$behav == "nestsearch", ]), 
      " overall nest searching with substantial mites")
paste(nrow(fq[fq$mites == "1" & fq$behav == "pollen", ]), 
      " overall pollen collecting with substantial mites")
paste(nrow(fq[fq$mites == "1" & fq$behav == "nectaring", ]), 
      " overall nectaring with substantial mites")

# 1-5 mites
paste(nrow(fq[fq$mitesAmount == "one" & fq$behav == "nestsearch", ]) + 
      nrow(fq[fq$mitesAmount == "two" & fq$behav == "nestsearch", ]) +
      nrow(fq[fq$mitesAmount == "three" & fq$behav == "nestsearch", ]) +
      nrow(fq[fq$mitesAmount == "four" & fq$behav == "nestsearch", ]) +
      nrow(fq[fq$mitesAmount == "five" & fq$behav == "nestsearch", ]), 
      " overall nest searching with 1-5 mites")
paste(nrow(fq[fq$mitesAmount == "one" & fq$behav == "pollen", ]) + 
      nrow(fq[fq$mitesAmount == "two" & fq$behav == "pollen", ]) +
      nrow(fq[fq$mitesAmount == "three" & fq$behav == "pollen", ]) +
      nrow(fq[fq$mitesAmount == "four" & fq$behav == "pollen", ]) +
      nrow(fq[fq$mitesAmount == "five" & fq$behav == "pollen", ]), 
      " overall pollen collecting with 5 or fewer mites")
paste(nrow(fq[fq$mitesAmount == "one" & fq$behav == "nectaring", ]) + 
      nrow(fq[fq$mitesAmount == "two" & fq$behav == "nectaring", ]) +
      nrow(fq[fq$mitesAmount == "three" & fq$behav == "nectaring", ]) +
      nrow(fq[fq$mitesAmount == "four" & fq$behav == "nectaring", ]) +
      nrow(fq[fq$mitesAmount == "five" & fq$behav == "nectaring", ]), 
      " overall nectaring with 1-5 mites")

# sphaer
fq[fq$sphaer == "1" & fq$behav == "nestsearch", ]
fq[fq$sphaer == "1" & fq$behav == "pollen", ]
fq[fq$sphaer == "1" & fq$behav == "nectaring", ]

# nosema
fq[fq$nosema == "1" & fq$behav == "nestsearch", ]
fq[fq$nosema == "1" & fq$behav == "pollen", ]
fq[fq$nosema == "1" & fq$behav == "nectaring", ]
```

# plot data
```{r}
# split out data, order behavior levels

foCA <- fo[startsWith(fo$beeID, "CA"),]
foME <- fo[startsWith(fo$beeID, "IME"),]
fqCA <- fq[startsWith(fo$beeID, "CA"),]
fqME <- fq[startsWith(fo$beeID, "IME"),]

foCA$behav <- ordered(foCA$behav, 
                      levels = c("nestsearch", "pollen", "nectaring"))
foME$behav <- ordered(foME$behav, 
                      levels = c("nestsearch", "pollen", "nectaring"))
fqCA$behav <- ordered(fqCA$behav, 
                      levels = c("nestsearch", "pollen", "nectaring"))
fqME$behav <- ordered(fqME$behav, 
                      levels = c("nestsearch", "pollen", "nectaring"))
fo$behav <- ordered(fo$behav, levels = c("nestsearch", "pollen", "nectaring"))
fq$behav <- ordered(fq$behav, levels = c("nestsearch", "pollen", "nectaring"))
```

## all species
```{r}
p1 <- ggplot(fo, aes(x = species, y = oocLen_unR, color = behav, group = behav)) +
  geom_hline(yintercept = mean(fo[fo$behav == "nestsearch", "oocLen_unR"], 
                               na.rm = TRUE),
             color="darkorchid4", alpha = 0.4) +
  geom_hline(yintercept = mean(fo[fo$behav == "pollen", "oocLen_unR"], 
                               na.rm = TRUE),
             color="cyan4", alpha = 0.4) +
  geom_hline(yintercept = mean(fo[fo$behav == "nectaring", "oocLen_unR"], 
                               na.rm = TRUE), 
             color="orange", alpha = 0.4) +
  geom_point(aes(shape = interaction(mites, sphaer)), 
             position = position_jitterdodge(jitter.width=0.6, 
                                             jitter.height=0.05, 
                                             dodge.width = 0.9),
             alpha = 0.7) +
  scale_shape_manual(values = c(21, 16, 2, 17), name = NULL, 
                     labels = c("0 mites,\n- Sphaerularia", 
                                ">5 mites,\n- Sphaerularia", 
                                "0 mites,\n+ Sphaerularia", 
                                ">5 mites,\n+ Sphaerularia")) +
  guides(shape = guide_legend(override.aes = list(size = 3))) +
  scale_color_manual(values = c("darkorchid4", "cyan4", "orange"), 
                     name = NULL, labels = c("Nest-searching", 
                                             "Pollen-collecting", 
                                             "Nectaring")) +
  guides(color = guide_legend(override.aes = list(shape = 15, size = 5))) +
  ylab("Oocyte length (mm)") +
  xlab("Species") +
  theme(text = element_text(size=12), axis.text.x = element_text(angle = 90), 
        legend.position = "bottom", legend.box="vertical")
p1

# oocyte resorption by species
p2 <- ggplot(fo[!duplicated(fo$beeID),], aes(x = species, y = oocPropR, 
                                             color = behav, group = behav)) +
  geom_hline(yintercept = mean(fo[fo$behav == "nestsearch", "oocPropR"], 
                               na.rm = TRUE),
             color="darkorchid4", alpha = 0.4) +
  geom_hline(yintercept = mean(fo[fo$behav == "pollen", "oocPropR"], 
                               na.rm = TRUE), 
             color="cyan4", alpha = 0.4) +
  geom_hline(yintercept = mean(fo[fo$behav == "nectaring", "oocPropR"], 
                               na.rm = TRUE), 
             color="orange", alpha = 0.4) +
  geom_point(aes(shape = interaction(mites, sphaer), color = behav), 
             position = position_jitterdodge(jitter.width=0.6, 
                                             jitter.height=0.05, 
                                             dodge.width = 0.9),
             alpha = 0.7) +
  scale_shape_manual(values = c(21, 16, 2, 17), name = NULL, 
                     labels = c("0 mites,\n- Sphaerularia", 
                                ">5 mites,\n- Sphaerularia", 
                                "0 mites,\n+ Sphaerularia", 
                                ">5 mites,\n+ Sphaerularia")) +
  guides(shape = guide_legend(override.aes = list(size = 3))) +
  scale_color_manual(values = c("darkorchid4", "cyan4", "orange"), 
                     name = NULL, labels = c("Nest-searching", 
                                             "Pollen-collecting", 
                                             "Nectaring")) + 
  guides(color = guide_legend(override.aes = list(shape = 15, size = 5))) +
  ylab("Proportion of oocytes resorbed") +
  xlab("Species") +
  theme(text = element_text(size=12), axis.text.x = element_text(angle = 90), 
        legend.position = "bottom", legend.box="vertical")

p2

# 2 part plot by behavior (all species)
ggarrange(p1, p2, nrow = 2, labels = c("A", "B"), 
          vjust = c(1.5, 0.3), 
          common.legend = TRUE, 
          legend = "bottom")
ggsave("figS2.jpg")
```


## vosnesenskii
```{r}
fo_sp <- fo[fo$species == "vosnesenskii",]
fq_sp <- fq[fq$species == "vosnesenskii",]

# max oocyte length by behavior (excluding resorbed oocytes)
ggplot(fq_sp, aes(behav, oocMaxUnR_Len, shape = mites, color = behav)) +
  geom_point(position = position_jitter(width=0.4, height=0.05), 
             alpha = 0.8, size = 3) +
  scale_shape_manual(values = c(21, 16), name = "Mites", labels = c("0", ">5")) +
  scale_color_manual(values = c("darkorchid4", "cyan4", "orange"), 
                     name = "Behavior", 
                     labels = c("Nest-searching", 
                                "Pollen-collecting", 
                                "Nectaring")) + 
  stat_summary(fun = mean, geom = "point", size = 6, shape = 15) +
  ylab("Oocyte length (mm)") +
  xlab("Behavior") +
  scale_x_discrete(labels=c("nestsearch" = "Nest-searching", 
                            "pollen" = "Pollen-collecting",
                            "nectaring" = "Nectaring")) +
  theme(text = element_text(size=14))

# oocyte length by behavior (excluding resorbed oocytes)
p1 <- ggplot(fo_sp, aes(behav, oocLen_unR, shape = mites, color = behav)) +
  geom_point(position = position_jitter(width=0.4, height=0.05), 
             alpha = 0.8, size = 3) +
  scale_shape_manual(values = c(21, 16), 
                     name = "Mites", 
                     labels = c("0", ">5")) +
  scale_color_manual(values = c("darkorchid4", "cyan4", "orange"), 
                     name = "Behavior", 
                     labels = c("Nest-searching", 
                                "Pollen-collecting", 
                                "Nectaring")) + 
  stat_summary(fun = mean, geom = "point", size = 6, shape = 15) +
  ylab("Oocyte length (mm)") +
  xlab("Behavior") +
  scale_x_discrete(labels=c("nestsearch" = "Nest-searching", 
                            "pollen" = "Pollen-collecting",
                            "nectaring" = "Nectaring")) +
  theme(text = element_text(size=14))

p1

# oocyte resorption by behavior
p2 <- ggplot(fq_sp, aes(behav, oocPropR, color = behav, shape = mites)) +
  geom_point(position = position_jitter(width=0.4, height=0.05), 
             alpha = 0.8, size = 3) +
  scale_shape_manual(values = c(21, 16), name = "Mites", 
                     labels = c("0", ">5")) +
  scale_color_manual(values = c("darkorchid4", "cyan4", "orange"), 
                     name = "Behavior", 
                     labels = c("Nest-searching", 
                                "Pollen-collecting", 
                                "Nectaring")) + 
  stat_summary(fun = mean, geom = "point", size = 6, shape = 15) +
  ylab("Proportion of oocytes resorbed") +
  xlab("Behavior") +
  scale_x_discrete(labels=c("nestsearch" = "Nest-searching", 
                            "pollen" = "Pollen-collecting",
                            "nectaring" = "Nectaring")) +
  theme(text = element_text(size=14))
p2

# oocyte lengths over time (excluding resorbed oocytes)
timeA <- ggplot(fo_sp, aes(date, oocLen_unR)) +
  geom_point(aes(shape = mites, color = behav),
             position = position_jitter(width=0.4, height=0.05), 
             alpha = 0.8, size = 2) +
  scale_shape_manual(values = c(21, 16), name = "Mites", 
                     labels = c("0", ">5")) + 
  ylab("Oocyte length (mm)") +
  xlab("Date") +
  scale_color_manual(values = c("darkorchid4", "cyan4", "orange"), 
                     name = "Behavior", 
                     labels = c("Nest-searching", 
                                "Pollen-collecting", 
                                "Nectaring")) +
  guides(color = guide_legend(override.aes = list(shape = 15, size = 5))) +
  theme(text = element_text(size=14))
timeA

# oocyte resorption over time (excluding resorbed occytes)
timeB <- ggplot(fo_sp, aes(date, oocPropR)) +
  geom_smooth(method=lm, alpha = 0, color = "gray") +
  geom_point(aes(shape = mites, color = behav), 
             position = position_jitter(width=0.4, height=0.05), 
             alpha = 0.8, size = 2) +
  ylab("Proportion of oocytes resorbed") +
  xlab("Date") +
  scale_shape_manual(values = c(21, 16), name = "Mites", 
                     labels = c("0", ">5")) +
  scale_color_manual(values = c("darkorchid4", "cyan4", "orange"), 
                     name = "Behavior", 
                     labels = c("Nest-searching", 
                                "Pollen-collecting", 
                                "Nectaring")) +
  guides(color = guide_legend(override.aes = list(shape = 15, size = 5))) +
  theme(text = element_text(size=14))
timeB

# 2 part plot by behavior
ggarrange(p1, p2, nrow = 2, labels = c("A", "B"), 
          common.legend = TRUE, 
          legend = "right")
ggsave("fig3.jpg")

# 2 part plot over time
ggarrange(timeA, timeB, nrow = 2, labels = c("A", "B"), 
          common.legend = TRUE, 
          legend = "right")
ggsave("fig4.jpg")
```
## ternarius
```{r}
fo_sp <- fo[fo$species == "ternarius",]
fq_sp <- fq[fq$species == "ternarius",]

# oocyte length (excluding resorbed oocytes)
ggplot(fo_sp, aes(behav, oocLen_unR, shape = mites, color = sphaer)) +
  geom_point(position = position_jitter(width=0.4, height=0.05), alpha = 0.8) +
  scale_shape_manual(values = c(21, 16), name = "mites") +
  stat_summary(fun = mean, geom = "point", size = 5, shape = 16) +
  ggtitle("unresrobed oocytes by behavior") +
  ylab("oocyte length (mm)") +
  xlab("behavior") +
  theme(text = element_text(size=15))


# oocyte stage
ggplot(fo_sp, aes(behav, stage, shape = mites, color = sphaer)) +
  geom_point(position = position_jitter(width=0.4, height=0.3), alpha = 0.8) +
  scale_shape_manual(values = c(21, 16), name = "mites") +
  stat_summary(fun = mean, geom = "point", size = 5, shape = 16) +
  ggtitle("unresorbed oocyte stages") +
  ylab("oocyte stage") +
  xlab("behavior") +
  theme(text = element_text(size=10))


# oocyte resorption
ggplot(fq_sp, aes(behav, oocPropR, color = sphaer, shape = mites)) +
  geom_point(position = position_jitter(width=0.4, height=0.05), alpha = 0.8) +
  scale_shape_manual(values = c(21, 16), name = "mites") +
  stat_summary(fun = mean, geom = "point", size = 5, shape = 16) +
  ggtitle("oocyte resorption") +
  ylab("proportion of oocytes resorbed") +
  xlab("behavior") +
  theme(text = element_text(size=10))

```

# analyze data

## oocyte length (t sphaer)
what predicts oocyte length?
```{r}
fo$behav <- as.character(fo$behav)

ggdensity(fo$oocLen_unR)
qqp(fo$oocLen_unR, "norm")

######################################### VOS ALONE -- null model fits best

fo_vos <- fo[fo$species == "vosnesenskii",]

lenFull <- glmer(oocLen_unR ~ behav + date + mites + (1|county) + (1|beeID), 
                family = gaussian(link = identity),
                data = fo_vos)
lenNull <- glmer(oocLen_unR ~ 1 + (1|county) + (1|beeID), 
                family = gaussian(link = identity),
                data = fo_vos)
len1 <- glmer(oocLen_unR ~ behav + mites + (1|county) + (1|beeID), 
                family = gaussian(link = identity),
                data = fo_vos)
len2 <- glmer(oocLen_unR ~ behav + date + (1|county) + (1|beeID), 
                family = gaussian(link = identity),
                data = fo_vos)
len3 <- glmer(oocLen_unR ~ mites + date + (1|county) + (1|beeID), 
                family = gaussian(link = identity),
                data = fo_vos)
len4 <- glmer(oocLen_unR ~ mites + (1|county) + (1|beeID), 
                family = gaussian(link = identity),
                data = fo_vos)
len5 <- glmer(oocLen_unR ~ date + (1|county) + (1|beeID), 
                family = gaussian(link = identity),
                data = fo_vos)
len6 <- glmer(oocLen_unR ~ behav + (1|county) + (1|beeID), 
                family = gaussian(link = identity),
                data = fo_vos)

model.sel(lenFull, lenNull, len1, len2, len3, len4, len5, len6)

tab_model(lenNull)

ggdensity(residuals(lenNull))

######################################### TERNARIUS ONLY -- sphaer significant

fo_tern <- fo[fo$species == "ternarius",]

lenFull <- glmer(oocLen_unR ~ behav + sphaer + mites + (1|beeID), 
                family = gaussian(link = identity),
                data = fo_tern)
lenNull <- glmer(oocLen_unR ~ 1 + (1|beeID), 
                family = gaussian(link = identity),
                data = fo_tern)
len1 <- glmer(oocLen_unR ~ behav + mites + (1|beeID), 
                family = gaussian(link = identity),
                data = fo_tern)
len2 <- glmer(oocLen_unR ~ behav + sphaer + (1|beeID), 
                family = gaussian(link = identity),
                data = fo_tern)
len3 <- glmer(oocLen_unR ~ mites + sphaer + (1|beeID), 
                family = gaussian(link = identity),
                data = fo_tern)
len4 <- glmer(oocLen_unR ~ behav + (1|beeID), 
                family = gaussian(link = identity),
                data = fo_tern)
len5 <- glmer(oocLen_unR ~ mites + (1|beeID), 
                family = gaussian(link = identity),
                data = fo_tern)
len6 <- glmer(oocLen_unR ~ sphaer + (1|beeID), 
                family = gaussian(link = identity),
                data = fo_tern)
model.sel(lenFull, lenNull, len1, len2, len3, len4, len5, len6)

tab_model(len6)
summary(len6)

ggdensity(residuals(len6))
```

## oocyte length max (t sphaer)
what predicts maximum oocyte length?
```{r}
fq$behav <- as.character(fq$behav)

ggdensity(fq$oocMaxUnR_Len)

######################################### VOS ALONE -- null model fits best

fq_vos <- fq[fq$species == "vosnesenskii",]
  
  lenFull <- glmer(oocMaxUnR_Len ~ behav + mites + date + (1|county), 
                  family = gaussian(link = identity),
                  data = fq_vos) 
  lenNull <- glmer(oocMaxUnR_Len ~ 1 + (1|county), 
                  family = gaussian(link = identity),
                  data = fq_vos)
  len1 <- glmer(oocMaxUnR_Len ~ behav + mites + (1|county), 
                  family = gaussian(link = identity),
                  data = fq_vos)
  len2 <- glmer(oocMaxUnR_Len ~ behav + date + (1|county), 
                  family = gaussian(link = identity),
                  data = fq_vos)
  len3 <- glmer(oocMaxUnR_Len ~ date + mites + (1|county), 
                  family = gaussian(link = identity),
                  data = fq_vos)
  len4 <- glmer(oocMaxUnR_Len ~ date + (1|county), 
                  family = gaussian(link = identity),
                  data = fq_vos)
  len5 <- glmer(oocMaxUnR_Len ~ behav + (1|county), 
                  family = gaussian(link = identity),
                  data = fq_vos)
  len6 <- glmer(oocMaxUnR_Len ~ mites + (1|county), 
                  family = gaussian(link = identity),
                  data = fq_vos)
    
model.sel(lenFull, lenNull, len1, len2, len3, len4, len5, len6)

tab_model(lenNull)

ggdensity(residuals(lenNull))


######################################### TERNARIUS ALONE -- sphaer significant

fq_tern <- fq[fq$species == "ternarius",]

lenFull <- glm(oocMaxUnR_Len ~ behav + sphaer + mites, 
                family = gaussian(link = identity),
                data = fq_tern)
lenNull <- glm(oocMaxUnR_Len ~ 1, 
                family = gaussian(link = identity),
                data = fq_tern)
len1 <- glm(oocMaxUnR_Len ~ behav + mites, 
                family = gaussian(link = identity),
                data = fq_tern)
len2 <- glm(oocMaxUnR_Len ~ mites + sphaer, 
                family = gaussian(link = identity),
                data = fq_tern)
len3 <- glm(oocMaxUnR_Len ~ sphaer + behav, 
                family = gaussian(link = identity),
                data = fq_tern)
len4 <- glm(oocMaxUnR_Len ~ behav, 
                family = gaussian(link = identity),
                data = fq_tern)
len5 <- glm(oocMaxUnR_Len ~ mites, 
                family = gaussian(link = identity),
                data = fq_tern)
len6 <- glm(oocMaxUnR_Len ~ sphaer, 
                family = gaussian(link = identity),
                data = fq_tern)
model.sel(lenFull, lenNull, len1, len2, len3, len4, len5, len6)

tab_model(len6)

ggdensity(residuals(len6))

```

## oocyte resorption (v date, t sphaer)
what predicts oocyte resorption?
```{r}
ggdensity(fq$oocPropR)

######################################### VOS ALONE -- date significant

fq_vos <- fq[fq$species == "vosnesenskii",]

resFull <- glmer(oocPropR ~ behav + mites + date + (1|county), 
                family = gaussian(link = identity),
                data = fq_vos)
resNull <- glmer(oocPropR ~ 1 + (1|county), 
                family = gaussian(link = identity),
                data = fq_vos)
res1 <- glmer(oocPropR ~ behav + date + (1|county), 
                family = gaussian(link = identity),
                data = fq_vos) 
res2 <- glmer(oocPropR ~ behav + mites + (1|county), 
                family = gaussian(link = identity),
                data = fq_vos)
res3 <- glmer(oocPropR ~ date + mites + (1|county), 
                family = gaussian(link = identity),
                data = fq_vos)
res4 <- glmer(oocPropR ~ date + (1|county), 
                family = gaussian(link = identity),
                data = fq_vos)
res5 <- glmer(oocPropR ~ behav + (1|county), 
                family = gaussian(link = identity),
                data = fq_vos)
res6 <- glmer(oocPropR ~ mites + (1|county), 
                family = gaussian(link = identity),
                data = fq_vos)

model.sel(resNull, resFull, res1, res2, res3, res4, res5, res6)

tab_model(res4)

ggdensity(residuals(res4))


######################################### TERNARIUS ONLY -- sphaer significant

fq_tern <- fq[fq$species == "ternarius",]

resFull <- glm(oocPropR ~ behav + sphaer + mites, 
                family = gaussian(link = identity),
                data = fq_tern)
resNull <- glm(oocPropR ~ 1, 
                family = gaussian(link = identity),
                data = fq_tern)
res1 <- glm(oocPropR ~ behav + mites, 
                family = gaussian(link = identity),
                data = fq_tern)
res2 <- glm(oocPropR ~ mites + sphaer, 
                family = gaussian(link = identity),
                data = fq_tern)
res3 <- glm(oocPropR ~ behav + sphaer, 
                family = gaussian(link = identity),
                data = fq_tern)
res4 <- glm(oocPropR ~ behav, 
                family = gaussian(link = identity),
                data = fq_tern)
res5 <- glm(oocPropR ~ mites, 
                family = gaussian(link = identity),
                data = fq_tern)
res6 <- glm(oocPropR ~ sphaer, 
                family = gaussian(link = identity),
                data = fq_tern)

model.sel(resFull, resNull, res1, res2, res3, res4, res5, res6)

tab_model(res3)
summary(res3)
ggdensity(residuals(res3))

```

## length - stage correlation
are oocyte length and stage correlated?
```{r}
coor <- lmer(oocLen_unR ~ stage + (1|beeID:bodySize),
    data = fo_vos)

coor <- lmer(oocLen_unR ~ stage + (1|species:beeID:bodySize),
    data = fo)

summary(coor)
tab_model(coor)

ggplot(fo_vos, aes(oocLen_unR, stage)) +
  geom_smooth(method=lm, alpha = 5, color = "gray20") +
  geom_point()
```


## mites
does mite presence correlate to behavioral category?
```{r}
chisq.test(x = fo_vos$behav, y = fo_vos$mites)
```
